---
title: "Q2"
author: "Anuja Janet"
date: "November 7, 2017"
output: html_document
---
```{r}
## Read the excel file.
ghg <- read.csv("D:\\college\\cs8086_data2decisions\\Project\\air-n-ghg.csv")
```
## Create a subset having CO2.
```{r}
ghg.sub <- ghg[ghg$SUBJECT == 'CO2', ]
```
## Create a subset with measure only Million tonnes.
```{r}
ghg.sub2 <- ghg.sub[ghg.sub$MEASURE == 'MLN_TONNE', ]
```
## Create a subset by removing WLD ans OED countries.
```{r}
ghg.sub3<-subset(ghg.sub2, ghg.sub2$COUNTRY!='WLD' & ghg.sub2$COUNTRY!='OED')
```
## Summary of the subset.
 ```{r}
summary(ghg.sub3)
```
## Create a basic scatter plot between Value and Urbanization to understand how they are correlated.
 ```{r}
 plot(ghg.sub3$Value~ghg.sub3$Urbanization.)
 ```
## Create another subset by removing outliers such as China and USA.
 ```{r}
ghg.sub4<-subset(ghg.sub3, ghg.sub3$ï..COUNTRY!='USA' & ghg.sub3$ï..COUNTRY!='CHN')
```
## Create a scatter plot between value and urbanization with the new subset that we have created.
 ```{r}
 plot(ghg.sub4$Value~ghg.sub4$Urbanization.)
 ```
## A best fit line enables us to understand the positive correlation better.
 ```{r}
abline(lm(ghg.sub4$Value~ghg.sub4$Urbanization.), col="red")
```
## Find the correlation value to confirm that it is positive.
 ```{r}
cor(ghg.sub4$Value, ghg.sub4$Urbanization) # =.22
```
 ```{r}
require(ggplot2)
```
## Graph CO2 and Urbanization percentage by country.
 ```{r}
ggplot(data = ghg.sub4, aes(x = Urbanization., y = Value, group = ï..COUNTRY)) +
xlab("Urbaniztion") + ylab("Millions of Tons") +
ggtitle("Total CO2 Emissions", subtitle="vs Urbanization") +
 geom_line(aes(color = ï..COUNTRY))
 ```
### These plots helps us confirm that as the urbanization percentage is increasing day-to day there is an increase in CO2 gas emmisions. We did not include CHINA and USA as they are outliers.
## This plot shows the correlation between Urbanization percentage and CO2 gas emmision with respect to each country including CHINA and USA.
 ```{r}
ggplot(data = ghg.sub3, aes(x = Urbanization., y = Value, group = ï..COUNTRY)) +
xlab("Urbaniztion") + ylab("Millions of Tons") +
ggtitle("Total CO2 Emissions", subtitle="vs Urbanization") +
 geom_line(aes(color = ï..COUNTRY))
 ```
